---
title: "Data_Types"
author: "Rebeccah Akinyi C029/401324/2023"
date: "`r Sys.Date()`"
output:
  word_document: default
  pdf_document: default
---

```{r setup2, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
```{r assignment, echo=TRUE}


# Data Types in R - Annotated Code Report
#In R, various data types are available to handle different kinds of data. Here is an overview with example, code and comments explaining each type.

## 1. Numeric
# Assign a numeric value to a variable
a<- 42 # 'a' stores a numeric value (double by default).

## 2. Integer
# Assign an integer value using the 'L' suffix
b<- 5L # 'b' is explicitly set as an integer.

## 3. Character (String)
# Assign a character string to a variable
char<- "Hello, World!" # 'char' is a string containing text.

## 4. Logical (Boolean)
# Assign a logical value (TRUE or FALSE)
c<- TRUE # 'c' holds a boolean value of TRUE.

## 5. Complex
# Create a complex number
s<- 4 + 3i # 's' holds a complex number (4 + 3i).

## 6. Vector
# Create a vector with different elements of the same type
v<- c(1, 2, 3, 4, 5) # 'v' is a numeric vector.

## 7. List
# Create a list that can contain different data types
p<- list(42, "text", TRUE, 3 + 2i) # 'p' contains multiple data types.

## 8. Factor
# Create a factor for categorical data
q<- factor(c("Low", "Medium", "High", "Medium")) # 'q' is a factor with levels.


## 9. Matrix
# Create a matrix with numeric data
m<- matrix(1:9, nrow=3, ncol=3) # 'mat' is a 3x3 numeric matrix.

## Summary
#Each of these examples demonstrates the basic data types in R. Understanding these data types is essential for working effectively with R.


