---
title: "Data Typesin R"
author: "Elly Ochieng'"
date: "`r Sys.Date()`"
output:
  word_document: default
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
```{r assignment, echo=TRUE}

# Data Types in R

# R treats everything as an object, and the simplest data objects are known as atomic data types.

# Atomic data types allow the creation of atomic vectors.
# Common atomic data types include:
# - Numeric (integer and double)
# - Character
# - Logical
# - Complex
# - Raw

# Checking if an object is atomic
is.atomic(3)            # TRUE, 3 is a numeric atomic vector
is.atomic("R CODER")    # TRUE, "R CODER" is a character atomic vector

# Functions to check data type in R
typeof(1)               # "double", shows the internal type
class(2)                # "numeric", shows the object's class
storage.mode(3)        # "double", shows storage mode
mode(4)                 # "numeric", another way to check type
str(5)                  # Displays the structure of the object, shows it is numeric

# Example of changing an object's class
x <- 1                  # Assigning numeric value 1
class(x)                # "numeric", initial class
class(x) <- "My_class"  # Changing class to "My_class"
class(x)                # "My_class", confirms class change
typeof(x)               # "double", still shows it's a double

# Summary of Type Functions
# Outputs of typeof, storage.mode, and mode for various types
# | Function       | logical | integer | double | character | raw  |
# |----------------|---------|---------|--------|-----------|------|
# | typeof         | logical | numeric | double | character | raw  |
# | storage.mode   | logical | numeric | double | character | raw  |
# | mode           | logical | integer | double | character | raw  |

# Numeric Data Types
# Numeric types consist of double and integer.
mode(55)                # "numeric"
is.numeric(3)          # TRUE, confirms it's numeric

# Double or Real Data Type
# Double-precision representation is default for all numbers
typeof(2)              # "double"
typeof(Inf)            # "double", shows infinity
typeof(-Inf)           # "double", shows negative infinity
typeof(NaN)            # "double", represents "Not a Number"
typeof(3.12e3)         # "double", shows scientific notation
typeof(0xbade)         # "double", hexadecimal notation

# Check if an object is double
is.double(2)          # TRUE
is.double(2.8)        # TRUE

# Integer Data Type
# Create integers by appending L to a number
y <- 2L                # Creates integer
typeof(y)              # "integer"
is.integer(3)          # FALSE, 3 is a double
is.integer(3L)         # TRUE, 3L is an integer

# Logical Data Type
# Composed of TRUE, FALSE, and NA
t <- TRUE              # Assign TRUE
f <- FALSE             # Assign FALSE
n <- NA                # Assign NA
typeof(t)              # "logical"
typeof(f)              # "logical"
typeof(n)              # "logical"

# Check if an object is logical
is.logical(T)          # TRUE
is.logical(TRUE)       # TRUE

# Caution with T and F
# Using T and F can override the values
F                     # FALSE, the default
a <- T                # Assigns TRUE to variable a
F <- a                # Now F is TRUE

# Complex Data Type
# Includes imaginary numbers
1 + 3i                # Represents a complex number
typeof(1 + 3i)        # "complex"
is.complex(1 + 3i)    # TRUE

# String or Character Data Type
# Character strings are enclosed in quotes
character <- "a"      # Assigns a string
typeof(character)     # "character"
is.character(character) # TRUE
typeof('R CODER')     # "character"
typeof("R CODER")     # "character"
nchar("A string")     # 8, counts characters including spaces

# Raw Data Type in R
# Holds raw bytes and is less common
a <- charToRaw("R CODER")  # Converts string to raw bytes                         # Outputs raw byte representation
typeof(a)                 # "raw"
b <- intToBits(3L)        # Converts integer to raw bits
typeof(b)                 # "raw"
is.raw(b)                # TRUE

# Date and Time Data Type in R
# Dates can be represented using as.Date()
date_example <- as.Date("2024-10-30")  # Example of creating a date object

# Data Types Coercion in R
# Coerce data types using functions that start with as.
# | Function        | Coerced Data Type |
# |-----------------|--------------------|
# | as.numeric      | Numeric            |
# | as.integer      | Integer            |
# | as.double       | Double             |
# | as.character    | Character          |
# | as.logical      | Boolean            |
# | as.raw          | Raw                |

# Example of coercing a double to integer
a <- 3                  # double by default
typeof(a)               # "double"
a <- as.integer(a)      # coerces to integer
typeof(a)               # "integer"

# Coerce logical values
b <- TRUE
b <- as.numeric(b)      # Coerces TRUE to 1
b                       # Outputs 1

c <- FALSE
c <- as.numeric(c)      # Coerces FALSE to 0
c                       # Outputs 0

d <- TRUE
d <- as.character(d)    # Coerces TRUE to string "TRUE"
d                       # Outputs "TRUE"

# Attempting to coerce incompatible types will yield an error
as.double("R CODER")    # Outputs NA with a warning: NAs introduced by coercion
