---
title: "Data Types in R"
author: "RISPER MUMBI MAINA"
date: "`r Sys.Date()`"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)  # or echo = FALSE
```
##R Markdown

##Data Types in R
```{r all-in-one, echo=TRUE, message=TRUE, warning=TRUE}
# Checking atomic data types
is.atomic(3)        # TRUE
is.atomic("R CODER") # TRUE

# Checking data types
typeof(1)      # "double"
class(2)       # "numeric"
storage.mode(3) # "double"
mode(4)        # "numeric"
str(5)         # num 5

# Numeric data types
typeof(2)      # "double"
typeof(Inf)    # "double"
typeof(-Inf)   # "double"
typeof(NaN)    # "double"
typeof(3.12e3) # "double"
typeof(0xbade) # "double"
is.double(2)   # TRUE
is.double(2.8) # TRUE

# Integer data type
y <- 2L
typeof(y)       # "integer"
is.integer(3)   # FALSE
is.integer(3L)  # TRUE

# Logical data type
t <- TRUE
f <- FALSE
n <- NA
typeof(t)    # "logical"
typeof(f)    # "logical"
typeof(n)    # "logical"
is.logical(T) # TRUE
is.logical(TRUE) # TRUE

# Complex data type
z <- 1 + 3i
typeof(z)      # "complex"
is.complex(z)  # TRUE

# Character data type
character <- "a"
typeof(character)      # "character"
is.character(character) # TRUE
typeof('R CODER')      # "character"
typeof("R CODER")      # "character"
nchar("A string")      # 8

# Raw data type
a <- charToRaw("R CODER")
typeof(a)     # "raw"
b <- intToBits(3L)
typeof(b)     # "raw"
is.raw(b)     # TRUE

# Date and Time data type
date <- as.Date("2023-10-30")
format(date, "%A %B %d %Y")

# Data type coercion
a <- 3
typeof(a)    # "double"
a <- as.integer(a)
typeof(a)    # "integer"

b <- TRUE
b <- as.numeric(b)
b            # 1

c <- FALSE
c <- as.numeric(c)
c            # 0

d <- TRUE
d <- as.character(d)
d            # "TRUE"

# Invalid coercion example
as.double("R CODER")  # Outputs: NA

